---
title: "Lab 11 - Grading the professor, Pt. 2"
author: Hannah Crawley
date: 4/7/2025
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```
# Part 1
## Exercise 1

```{r exercise1}
m_bty <- linear_reg() %>%
  set_engine("lm") %>%
  fit(score ~ bty_avg, data = evals)
```

# Part 2
## Exercise 2
> Fit a linear model (one you have fit before): m_bty_gen, predicting average professor evaluation score based on average beauty rating (bty_avg) and gender. Write the linear model, and note the R2 and the adjusted R2.

 
```{r exercise2_code}
```

## Exercise 3
> Interpret the slope and intercept of m_bty_gen in context of the data. 

## Exercise 4
> What percent of the variability in score is explained by the model m_bty_gen.


## Exercise 5
> What is the equation of the line corresponding to just male professors?

## Exercise 6
> For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

## Exercise 7
> How does the relationship between beauty and evaluation score vary between male and female professors?

## Exercise 8 
>How do the adjusted R2 values of m_bty_gen and m_bty compare? What does this tell us about how useful gender is in explaining the variability in evaluation scores when we already have information on the beauty score of the professor. 

## Exercise 9
> Compare the slopes of bty_avg under the two models (m_bty and m_bty_gen). Has the addition of gender to the model changed the parameter estimate (slope) for bty_avg?

## Exercise 10
> Create a new model called m_bty_rank with gender removed and rank added in. Write the equation of the linear model and interpret the slopes and intercept in context of the data.

# Part 3: The search for the best model 
Going forward, only consider the following variables as potential predictors: rank, ethnicity, gender, language, age, cls_perc_eval, cls_did_eval, cls_students, cls_level, cls_profs, cls_credits, bty_avg.

## Exercise 11
> Which variable, on its own, would you expect to be the worst predictor of evaluation scores? Why? Hint: Think about which variable would you expect to not have any association with the professor’s score.

## Exercise 12
> Check your suspicions from the previous exercise. Include the model output for that variable in your response.

## Exercise 13
> Suppose you wanted to fit a full model with the variables listed above. If you are already going to include cls_perc_eval and cls_students, which variable should you not include as an additional predictor? Why?

## Exercise 14
> Fit a full model with all predictors listed above (except for the one you decided to exclude) in the previous question.

## Exercise 15
> Using backward-selection with adjusted R-squared as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on.

## Exercise 16
> Interpret the slopes of one numerical and one categorical predictor based on your final model.

## Exercise 17
> Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.

## Exercise 18
> Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?